---
title: "David Lovelace Archive — Project Bid"
subtitle: "Progressing the archive: stakeholder liaison, georectification, web development, and outreach"
author: "Robin Lovelace"
date: last-modified
format:
  gfm:
    toc: true
    toc-title: Contents
execute:
  eval: true
  echo: false
  message: false
  warning: false
---

```{r}
# | label: setup
# | include: false
library(knitr)
```

## Summary

This project will progress the David Lovelace Archive (DLA) from a static
prototype to an actively managed, openly accessible resource that serves
Herefordshire's conservation, heritage, and research communities. The project
has four workstreams, designed to create lasting value for the county, to
improve knowledge of the valuable information in the archive, and ensuring
people across the county and beyond can benefit from this unique record.

The David Lovelace Archive is a substantial collection of maps, aerial
photography, habitat surveys, veteran tree records, and historical documents
assembled over decades by David Lovelace (1948–2026). It represents the most
comprehensive non-governmental archive of Herefordshire's landscape and
ecological history. This bid seeks to make it accessible, geospatially
enabled, and actively used by the organisations and communities researching
and preserving.

**Timescale:** 12 months.

## Organisations and context

- **Herefordshire Meadows** — 5 days. Benefit from georectification of meadow
  sites, veteran tree survey data, QGIS training, and open access to archive
  information.

- **Herefordshire Wildlife Trust** — 2 days. Benefit from habitat survey data,
  nature reserve boundary overlays, and training in using historic maps for
  landscape recovery.

- **Woolhope Club (Woolhope Naturalists' Field Club)** — part of Robin
  Lovelace's project lead time. Benefit from archive transfer to their library,
  access to tithe maps and survey photographs for their 600+ members.

- **Herefordshire Archive and Records Centre (HARC)** — part of Robin
  Lovelace's project lead time. Benefit from transfer of priority digital and
  physical materials, ensuring long-term public access and integration with
  existing collections.

- **Ancient Tree Forum (ATF)** — 5 days. Benefit from the LOWVP Veteran Tree
  Survey being made openly accessible and cross-referenced with the Ancient
  Tree Inventory, ground-truthing field work, and training.

- **Homme House, Much Marcle** — 2 days (in-kind). Benefit from inclusion as a
  case study site, georectification of estate records, parkland tree data
  digitisation, and training opportunities.

- **Stuart Hedley — Botanical Consultant** — 5 days. Benefit from access to
  the archive's botanical and habitat records for VC36 recording, training in
  QGIS for field botany, and integration with Herefordshire Meadows work.

---

## Workstreams

### Workstream 1: Stakeholder liaison and archive prioritisation

```{r}
# | label: ws1-days
ws1_robin <- 4
ws1_stuart <- 5
ws1_atf <- 2
ws1_meadows <- 2
ws1_wildlife <- 1
ws1_homme <- 1

ws1 <- data.frame(
  Person = c("Robin Lovelace", "Stuart Hedley", "Ancient Tree Forum",
             "Herefordshire Meadows", "Herefordshire Wildlife Trust", "Homme House"),
  Days = c(ws1_robin, ws1_stuart, ws1_atf, ws1_meadows, ws1_wildlife, ws1_homme),
  Role = c(
    "Project lead; stakeholder meetings with HARC, Woolhope Club, others",
    "Botanical audit; identify priority meadows and woodland surveys",
    "Audit veteran tree survey data; identify priority parishes for release",
    "Identify meadow restoration sites with matching archival records",
    "Identify nature reserve overlap with archive habitat data",
    "Identify veteran parkland trees and estate records for inclusion"
  )
)
kable(ws1, caption = "Workstream 1 — Stakeholder liaison and prioritisation effort")
```

**Activities:**

- Initial meetings with each stakeholder to understand their priorities and
  data needs.

- Audit of archive contents relevant to each organisation's conservation and
  research remit.

- Prioritisation matrix — conservation value, archival completeness, and
  stakeholder demand.

- Agreed publication schedule for data release.

- Transfer of priority digital and physical materials to HARC, the Woolhope
  Club library, and other partner organisations for long-term public access.

---

### Workstream 2: Georectification of historic maps and aerial photography

```{r}
# | label: ws2-days
ws2_robin <- 4
ws2_stuart <- 0
ws2_atf <- 2
ws2_meadows <- 2
ws2_wildlife <- 1
ws2_homme <- 1

ws2 <- data.frame(
  Person = c("Robin Lovelace", "Stuart Hedley", "Ancient Tree Forum",
             "Herefordshire Meadows", "Herefordshire Wildlife Trust", "Homme House"),
  Days = c(ws2_robin, ws2_stuart, ws2_atf, ws2_meadows, ws2_wildlife, ws2_homme),
  Role = c(
    "Lead georectification; QGIS processing; PMTiles pipeline",
    "—",
    "Ground-truth veteran tree locations; verify historic grid references",
    "Provide meadow boundary data for historic map overlay",
    "Provide nature reserve boundary data",
    "Provide parkland boundary and veteran tree location data"
  )
)
kable(ws2, caption = "Workstream 2 — Georectification effort")
```

**Activities:**

- Georectify priority tithe maps (Kings Caple, Clifford, selected parishes).

- Georectify RAF 1947 and Luftwaffe aerial photography covering priority areas.

- Create PMTiles and COG (Cloud Optimised GeoTIFF) derivatives for web
  deployment.

- Ground-truth selected historic features with stakeholder partners.

- Publish as browsable, interactive map layers on the archive website.

- Cross-reference veteran tree survey photos with the Ancient Tree Inventory
  and make available via ATF's national network.

---

### Workstream 3: Outreach, QGIS training, and research initiation

```{r}
# | label: ws3-days
ws3_robin <- 3
ws3_stuart <- 0
ws3_atf <- 1
ws3_meadows <- 1
ws3_wildlife <- 0
ws3_homme <- 0

ws3 <- data.frame(
  Person = c("Robin Lovelace", "Stuart Hedley", "Ancient Tree Forum",
             "Herefordshire Meadows", "Herefordshire Wildlife Trust", "Homme House"),
  Days = c(ws3_robin, ws3_stuart, ws3_atf, ws3_meadows, ws3_wildlife, ws3_homme),
  Role = c(
    "Design and deliver QGIS training; academic liaison; workshop coordination",
    "—",
    "Contribute veteran tree survey methods to workshop content",
    "Co-host meadow-focused session; engage network of land managers",
    "—",
    "—"
  )
)
kable(ws3, caption = "Workstream 3 — Outreach and research effort")
```

**Activities:**

- **David Lovelace Archive Workshop** — a 1–2 day hands-on training event
  designed for complete beginners, introducing QGIS as a tool for exploring
  historic maps, aerial photography, and ecological survey data.

- Workshop participants will learn to load georectified maps, overlay modern
  boundaries, digitise features from historic sources, and capture field data
  using mobile QGIS.

- Stakeholders (Herefordshire Meadows, ATF, Woolhope Club, Homme House) will
  contribute use cases drawn from their own sites and datasets.

- Academic liaison with the University of Leeds (Institute for Transport
  Studies) and other research partners to identify at least one research
  project using the archive's data — for example, measuring change in
  veteran tree populations, historic field boundary loss, or woodland
  continuity over 150+ years.

- Workshop materials (data packs, QGIS project files, video guides) will be
  published openly for reuse.

---

### Workstream 4: Web development — asset geolocation and hosting

```{r}
# | label: ws4-days
ws4_robin <- 4

ws4 <- data.frame(
  Person = c("Robin Lovelace"),
  Days = c(ws4_robin),
  Role = c("Lead developer; geolocation feature implementation; deployment")
)
kable(ws4, caption = "Workstream 4 — Web development effort")
```

**Activities:**

- Add **asset geolocation functionality** to the archive website: a browsable,
  searchable map interface that displays the location of every photograph, map,
  survey, and document in the archive, with spatial filtering by parish, grid
  square, and stakeholder area of interest.

- Develop a **hosting pipeline** for the PMTiles, COG, and other spatial data
  layers created in Workstream 2, ensuring they're available for web
  publication at no ongoing cost (e.g. via free static hosting on Netlify,
  GitHub Pages, or similar).

- Integrate a **download portal** for bulk data access, so researchers and
  partner organisations can access the archive's content without manual file
  requests.

- Ensure the site architecture supports long-term sustainability: the openclaw
  server currently houses the clean archive (~2.4 TB), and the web interface
  will need to point to data published via GitHub Releases and static hosting.

---

## Resource summary

```{r}
# | label: resource-summary
totals <- data.frame(
  Person = c("Robin Lovelace", "Stuart Hedley", "Ancient Tree Forum",
             "Herefordshire Meadows", "Homme House", "Herefordshire Wildlife Trust"),
  WS1 = c(ws1_robin, ws1_stuart, ws1_atf, ws1_meadows, ws1_homme, ws1_wildlife),
  WS2 = c(ws2_robin, ws2_stuart, ws2_atf, ws2_meadows, ws2_homme, ws2_wildlife),
  WS3 = c(ws3_robin, ws3_stuart, ws3_atf, ws3_meadows, ws3_homme, ws3_wildlife),
  WS4 = c(ws4_robin, 0, 0, 0, 0, 0),
  Total = c(
    ws1_robin + ws2_robin + ws3_robin + ws4_robin,
    ws1_stuart + ws2_stuart + ws3_stuart,
    ws1_atf + ws2_atf + ws3_atf,
    ws1_meadows + ws2_meadows + ws3_meadows,
    ws1_homme + ws2_homme + ws3_homme,
    ws1_wildlife + ws2_wildlife + ws3_wildlife
  )
)
kable(totals, caption = "Resource allocation by person and workstream (days)")

total_days <- sum(totals$Total)
```

**Total person-days: `r total_days`**

| Person | Days | Notes |
|--------|------|-------|
| Robin Lovelace | `r totals$Total[totals$Person == "Robin Lovelace"]` | Project lead, georectification, web dev, training |
| Stuart Hedley | `r totals$Total[totals$Person == "Stuart Hedley"]` | Botanical audit and survey |
| Ancient Tree Forum | `r totals$Total[totals$Person == "Ancient Tree Forum"]` | Veteran tree expertise |
| Herefordshire Meadows | `r totals$Total[totals$Person == "Herefordshire Meadows"]` | Meadow conservation and data |
| Homme House | `r totals$Total[totals$Person == "Homme House"]` | In-kind: parkland records, venue, case study |
| Herefordshire Wildlife Trust | `r totals$Total[totals$Person == "Herefordshire Wildlife Trust"]` | Reserve data and habitat records |
| **Total** | **`r total_days`** | |

---

## Timeline (12 months)

| Quarter | Workstream focus | Key milestones |
|---------|-----------------|----------------|
| Q1 (Months 1–3) | WS1: Stakeholder liaison | Initial meetings, archive audit, prioritisation, transfer agreements signed |
| Q2 (Months 4–6) | WS2: Georectification | Priority tithe maps and aerial photos georectified; first PMTiles published |
| Q3 (Months 7–9) | WS3: Outreach + WS4: Web dev | QGIS workshop delivered; asset geolocation feature live; data transferred to HARC |
| Q4 (Months 10–12) | All | Cross-referencing complete; materials transferred to Woolhope Club and partners; project report |

---

## Proposed budget

| Item | Days | Day rate (£) | Total (£) |
|------|------|-------------|-----------|
| Robin Lovelace | `r ws1_robin+ws2_robin+ws3_robin+ws4_robin` | TBC | TBC |
| Stuart Hedley | `r ws1_stuart+ws2_stuart+ws3_stuart` | TBC | TBC |
| Ancient Tree Forum | `r ws1_atf+ws2_atf+ws3_atf` | TBC | TBC |
| Herefordshire Meadows | `r ws1_meadows+ws2_meadows+ws3_meadows` | TBC | TBC |
| Herefordshire Wildlife Trust | `r ws1_wildlife+ws2_wildlife+ws3_wildlife` | TBC | TBC |
| Homme House | 2 days in-kind | — | — |
| **Total** | **`r total_days`** | | **TBC** |

---

## Next steps

1. Confirm stakeholder interest, availability, and in-kind commitments.

2. Finalise day rates and budget figures.

3. Submit to Brightspace Foundation (first instance).

4. On award: recruit project team, schedule initial stakeholder roundtable,
   and begin archive audit.
