---
title: "David Lovelace Archive — Project Bid"
subtitle: "Progressing the archive: stakeholder liaison, georectification, web development, and outreach"
author: "Robin Lovelace"
date: last-modified
format:
  gfm:
    toc: true
    toc-title: Contents
execute:
  eval: true
  echo: false
  message: false
  warning: false
---

```{r}
# | label: setup
# | include: false
library(knitr)
```

## Summary

This project will progress the David Lovelace Archive (DLA) from a static
prototype to an actively managed, openly accessible resource that serves
Herefordshire's conservation, heritage, and research communities. The project
has four workstreams, designed to create lasting value for the county, to
improve knowledge of the valuable information in the archive, and ensuring
people across the county and beyond can benefit from this unique record.

The David Lovelace Archive is a substantial collection of maps, aerial
photography, habitat surveys, veteran tree records, and historical documents
assembled over decades by David Lovelace (1948–2026). It represents the most
comprehensive non-governmental archive of Herefordshire's landscape and
ecological history. This bid seeks to make it accessible, geospatially
enabled, and actively used by the organisations and communities researching
and preserving.

**Timescale:** 12 months.

## Organisations and context

- **Herefordshire Meadows** — 5 days. Georectification of meadow
  sites, veteran tree survey data, QGIS training, and open access to archive
  information.

- **Herefordshire Wildlife Trust** — 2 days. Habitat survey data,
  nature reserve boundary overlays, and training in using historic maps for
  landscape recovery.

- **Woolhope Club (Woolhope Naturalists' Field Club)** — part of Robin
  Lovelace's project lead time. Archive transfer to their library,
  access to tithe maps and survey photographs for their 600+ members.

- **Herefordshire Archive and Records Centre (HARC)** — part of Robin
  Lovelace's project lead time. Transfer of priority digital and
  physical materials, ensuring long-term public access and integration with
  existing collections.

- **Ancient Tree Forum (ATF)** — 5 days. LOWVP Veteran Tree
  Survey made openly accessible and cross-referenced with the Ancient
  Tree Inventory, ground-truthing field work, and training.

- **Homme House, Much Marcle** — 2 days (in-kind). Inclusion as a
  case study site, georectification of estate records, parkland tree data
  digitisation, and training opportunities.

- **Stuart Hedley — Botanical Consultant** — 5 days. Access to
  the archive's botanical and habitat records for VC36 recording, training in
  QGIS for field botany, and integration with Herefordshire Meadows work.

---

## Workstreams

### Workstream 1: Stakeholder liaison and archive prioritisation

```{r}
# | label: ws1-days
ws1_robin <- 4
ws1_stuart <- 5
ws1_atf <- 2
ws1_meadows <- 2
ws1_wildlife <- 1
ws1_homme <- 1

ws1 <- data.frame(
  Person = c("Robin Lovelace", "Stuart Hedley", "Ancient Tree Forum",
             "Herefordshire Meadows", "Herefordshire Wildlife Trust", "Homme House"),
  Days = c(ws1_robin, ws1_stuart, ws1_atf, ws1_meadows, ws1_wildlife, ws1_homme),
  Role = c(
    "Project lead; stakeholder meetings with HARC, Woolhope Club, others",
    "Botanical audit; identify priority meadows and woodland surveys",
    "Audit veteran tree survey data; identify priority parishes for release",
    "Identify meadow restoration sites with matching archival records",
    "Identify nature reserve overlap with archive habitat data",
    "Identify veteran parkland trees and estate records for inclusion"
  )
)
kable(ws1, caption = "Workstream 1 — Stakeholder liaison and prioritisation effort")
```

**Activities:**

- Stakeholder meetings to agree priorities and data needs.

- Archive audit per stakeholder conservation and research remit.

- Prioritisation matrix, publication schedule, and materials transfer
  to HARC and Woolhope Club.

---

### Workstream 2: Georectification of historic maps and aerial photography

```{r}
# | label: ws2-days
ws2_robin <- 4
ws2_stuart <- 0
ws2_atf <- 2
ws2_meadows <- 2
ws2_wildlife <- 1
ws2_homme <- 1

ws2 <- data.frame(
  Person = c("Robin Lovelace", "Ancient Tree Forum",
             "Herefordshire Meadows", "Herefordshire Wildlife Trust", "Homme House"),
  Days = c(ws2_robin, ws2_atf, ws2_meadows, ws2_wildlife, ws2_homme),
  Role = c(
    "Lead georectification; QGIS processing; PMTiles pipeline",
    "Ground-truth veteran tree locations; verify historic grid references",
    "Provide meadow boundary data for historic map overlay",
    "Provide nature reserve boundary data",
    "Provide parkland boundary and veteran tree location data"
  )
)
kable(ws2, caption = "Workstream 2 — Georectification effort")
```

**Activities:**

- Georectify priority tithe maps (Kings Caple, Clifford, selected parishes).

- Georectify RAF 1947 and Luftwaffe aerial photography covering priority areas.

- Create PMTiles and COGs for web deployment.

- Ground-truth selected historic features with stakeholder partners.

- Publish as browsable, interactive map layers on the archive website.

- Cross-reference veteran tree survey with the Ancient Tree Inventory
  for ATF's national network.

---

### Workstream 3: Outreach, QGIS training, and research initiation

```{r}
# | label: ws3-days
ws3_robin <- 3
ws3_stuart <- 0
ws3_atf <- 1
ws3_meadows <- 1
ws3_wildlife <- 0
ws3_homme <- 0

ws3 <- data.frame(
  Person = c("Robin Lovelace", "Ancient Tree Forum",
             "Herefordshire Meadows"),
  Days = c(ws3_robin, ws3_atf, ws3_meadows),
  Role = c(
    "Design and deliver QGIS training; academic liaison; workshop coordination",
    "Contribute veteran tree survey methods to workshop content",
    "Co-host meadow-focused session; engage network of land managers"
  )
)
kable(ws3, caption = "Workstream 3 — Outreach and research effort")
```


**Activities:**

- **David Lovelace Archive Workshop** — 1–2 day QGIS beginner workshop:
  loading georectified maps, overlaying modern boundaries, digitising
  historic features, and mobile field data capture.

- Stakeholders (Herefordshire Meadows, ATF, Woolhope Club, Homme House)
  contribute use cases from their own sites and datasets.

- Academic liaison with University of Leeds ITS and other partners to
  identify at least one research project from the archive — measuring
  change in veteran tree populations, field boundary loss, or woodland
  continuity over 150+ years.

- Workshop materials (data packs, QGIS project files, video guides)
  published openly for reuse.

---

### Workstream 4: Web development — asset geolocation and hosting

```{r}
# | label: ws4-days
ws4_robin <- 4

ws4 <- data.frame(
  Person = c("Robin Lovelace"),
  Days = c(ws4_robin),
  Role = c("Lead developer; geolocation feature implementation; deployment")
)
kable(ws4, caption = "Workstream 4 — Web development effort")
```

**Activities:**

- **Asset geolocation** — browsable, searchable map interface with
  spatial filtering by parish, grid square, and stakeholder area of interest.

- **Hosting pipeline** for PMTiles and COG layers from WS2,
  published at no ongoing cost via Netlify, GitHub Pages, or similar.

- **Download portal** for bulk data access — researchers and partners
  can retrieve archive content without manual file requests.

- **Long-term sustainability** — clean archive (~2.4 TB on Leeds
  server) served via GitHub Releases and static hosting.

---

## Resource summary

```{r}
# | label: resource-summary
totals <- data.frame(
  Person = c("Robin Lovelace", "Stuart Hedley", "Ancient Tree Forum",
             "Herefordshire Meadows", "Homme House", "Herefordshire Wildlife Trust"),
  WS1 = c(ws1_robin, ws1_stuart, ws1_atf, ws1_meadows, ws1_homme, ws1_wildlife),
  WS2 = c(ws2_robin, ws2_stuart, ws2_atf, ws2_meadows, ws2_homme, ws2_wildlife),
  WS3 = c(ws3_robin, ws3_stuart, ws3_atf, ws3_meadows, ws3_homme, ws3_wildlife),
  WS4 = c(ws4_robin, 0, 0, 0, 0, 0),
  Total = c(
    ws1_robin + ws2_robin + ws3_robin + ws4_robin,
    ws1_stuart + ws2_stuart + ws3_stuart,
    ws1_atf + ws2_atf + ws3_atf,
    ws1_meadows + ws2_meadows + ws3_meadows,
    ws1_homme + ws2_homme + ws3_homme,
    ws1_wildlife + ws2_wildlife + ws3_wildlife
  )
)
kable(totals, caption = "Resource allocation by person and workstream (days)")

total_days <- sum(totals$Total)
```

**Total person-days: `r total_days`**

---

## Timeline (12 months)

| Quarter | Workstream focus | Key milestones |
|---------|-----------------|----------------|
| Q1 (Months 1–3) | WS1: Stakeholder liaison | Initial meetings, archive audit, prioritisation, transfer agreements signed |
| Q2 (Months 4–6) | WS2: Georectification | Priority tithe maps and aerial photos georectified; first PMTiles published |
| Q3 (Months 7–9) | WS3: Outreach + WS4: Web dev | QGIS workshop delivered; asset geolocation feature live; data transferred to HARC |
| Q4 (Months 10–12) | All | Cross-referencing complete; materials transferred to Woolhope Club and partners; project report |

---

## Proposed budget

| Item | Days | Day rate (£) | Total (£) |
|------|------|-------------|-----------|
| Robin Lovelace | `r ws1_robin+ws2_robin+ws3_robin+ws4_robin` | TBC | TBC |
| Stuart Hedley | `r ws1_stuart+ws2_stuart+ws3_stuart` | TBC | TBC |
| Ancient Tree Forum | `r ws1_atf+ws2_atf+ws3_atf` | TBC | TBC |
| Herefordshire Meadows | `r ws1_meadows+ws2_meadows+ws3_meadows` | TBC | TBC |
| Herefordshire Wildlife Trust | `r ws1_wildlife+ws2_wildlife+ws3_wildlife` | TBC | TBC |
| Homme House | 2 days in-kind | — | — |
| **Total** | **`r total_days`** | | **TBC** |

---

## Next steps

1. Confirm stakeholder interest, availability, and in-kind commitments.

2. Finalise day rates and budget figures.

3. Submit to Brightspace Foundation (first instance).

4. On award: recruit project team, schedule initial stakeholder roundtable,
   and begin archive audit.
